Galois Slicing as Differentiation

### 2025-07-11 main/conclusion.tex
[ tex ] Bit more summary text in conclusion
[ tex ] save a line

We have presented a semantic version of \GPS, inspired by connections to differentiable programming and
automatic differentation, and shown that it can be used to interpret an expressive higher-order language
suitable for writing simple queries and data manipulation. Our model elucidates some of the decisions
implicitly taken in previous works on \GPS (\secref{cbn-translation}), and reveals new applications such as
approximation by intervals \secref{interval-approx}. Our categorical approach admits a modular construction of
our model, and the use of general theorems, such as Fiore and Simpson's \thmref{glr-definability}, to prove
properties of the interpretation.

@@

\paragraph{Categorical Models of Differentation} We have already identified \conref{tangent-stable-fns} and
\conref{tangent-fam} as potential connections to categorical models of differentation. It would also be
interesting to discover any possible links to Differential Linear Logic
\cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
give an account of reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
of \citet{vakar22}.

>>>>

We have presented a semantic version of \GPS, inspired by connections to differentiable programming and
automatic differentation, and shown that it can be used to interpret an expressive higher-order language
suitable for writing simple queries and data manipulation. Our model elucidates some of the decisions
implicitly taken in previous works on \GPS (\secref{cbn-translation}), and reveals new applications such as
approximation by intervals \secref{interval-approx}. Our categorical approach admits a modular construction of
our model, and the use of general theorems, such as Fiore and Simpson's \thmref{glr-definability}, to prove
properties of the interpretation. We have focused on constructions that enable an executable implementation in
Agda in this work, but have conjectured that there are connections to established notions of categorical
differentiation in \conref{tangent-stable-fns} and \conref{tangent-fam}.

@@

\paragraph{Categorical Models of Differentation} We have already identified \conref{tangent-stable-fns} and
\conref{tangent-fam} as potential connections to categorical models of differentation. It would also be
interesting to discover any possible links to Differential Linear Logic
\cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
model reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
of \citet{vakar22}.
